### 2024-02-06 ms.qmd
Updates

Behavioural flexibility, the ability to adjust the behaviour adaptively in response to internal or external by, e.g., problem-solving, innovation, or alterations in one behaviour in favour of another [@brown2010behavioral], is considered crucial for invasion success [@wright2010behavioral; @chapple2012can; @szabo2020can]. In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats, as behaving flexibly enhances their capacity to exploit new resources and face new challenges [@wright2010behavioral; @chapple2012can; @szabo2020can]. Indeed, when closely related species of birds were compared, successful invaders showed a higher frequency of foraging innovations [@sol2000behavioural]. According to the adaptive flexibility hypothesis, behavioural flexibility can be more relevant during introduction, establishment, and spread of the population [@wright2010behavioral; @chapple2012can; @szabo2020can], stages that, nonetheless, can involve more than one generation. If behavioural flexibility is adaptive in these different stages, both this trait and the underlining mechanisms are expected to persist through generations in succesful invaders. For example, relative brain size is a good predictor of behavioural flexibility and invasive success in mammals [@sol2008brain], birds [@sol2000behavioural; @sol2002behavioural], reptiles, and amphibians [@amiel_effects_2017]. Since the brain is especially sensitive to environmental changes during the early stages of development [@zhu_prenatal_2004] it is possible that the new prenatal environment faced by establishing populations can affect the behavioural flexibility of the new generations [see @jonson1976effect; @lui2017chronic], with a consequent impact on invasion success. Therefore, we consider important to investigate the effect of developmental conditions on behavioural flexibility and whether their susceptibility to the early environment can predict the success of invasive species.  
In this scenario, we consider prenatal GCs and thermal environment to be salient factors in the study of behavioural flexibility within the context of invasion biology for several reasons. First, GCs are the main group of hormones related to animals’ response towards alterations in homeostasis [@sapolsky_how_2000] that may become relevant when facing new environments. Furthermore, Gcs are one of the mechanisms mediating vertebrates’ response to changes in temperature that, at the same time, can have direct effects on animals' physiology and behaviour [@Crino_2023]. Secondly, there is evidence that both developmental GCs and thermal environment can impact cognition in different taxa [@jonson1976effect; @zhu_prenatal_2004; @amiel_hotter_2012; @lui2017chronic]. For instance, young Sprague-Dawley rats treated with Corticosterone (CORT) - the principal GC in rodents, birds, and reptiles [@Crino_2023] - performed worse than controls in a reversal learning task [@lui2017chronic], while exposure of pregnant guinea pigs to higher temperatures resulted in significant impariments with offspring behavoural flexibility [@jonson1976effect]. In addition, the cognitive effects of both factors are triggered by similar physiological mechanisms intertwined with brain development [@jonson1976effect; @coomber_independent_1997; @zhu_prenatal_2004], an aspect that is considered essential in succesful invaders [@sol2000behavioural; @sol2002behavioural; @sol2008brain; @amiel_smart_2011].
Here, we studied the interactive contribution of prenatal temperature and Glucocorticoids (GCs) on behavioural flexibility in two species that exhibit distinct outcomes in their roles as invasive species, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L. delicata_ has been succesful in colonising different areas around the globe [@baker1979rainbow; @chapple2013divergent; @chapple2015deliinvLHI]. Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L. delicata_, and are likely key for its invassive success [@chapple_know_2011; @chapple2012can; @bezzina2014does]. However, differences between species in behavioural flexibility and how this trait is affected by the prenatal environment remains unknown. To investigate these effects, we first manipulated CORT levels and incubation temperature in eggs from both species, and then subjected the juveniles to a reversal learning task as a measure of behavioural flexibility [@brown2010behavioral; @szabo2020can]. We predicted that _L. delicata_ would, on average, perform more proficiently in the task that _L. guichenoti_. We also anticipated prenatal CORT and temperature to affect the behavioural flexibility of both species, and that _L. guichenoti_ would be more impacted by the prenatal environment than _L. delicata_. Regarding the directionality of the treatments, high prenatal temperatures are considered to improve performance in cognitive tasks in ectotherms [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014], so we expected those incubated at higher temperatures to perform better in the reversal task. On the other hand, we expected that prenatal CORT would impair the performance of the lizards in the task, as it has been shown in other species [@lemaire_prenatal_2000; @zhu_prenatal_2004; @lui2017chronic].

@@

We performed the analyses for both species separately. As such, we fitted a total of two different Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022). We had per each model four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations. We used 'Choice' - whether the individual chose correct (1) or not (0) - as the response variable, and trial, hormone ('CORT' versus 'Control'), incubation temperature ('Cold' versus 'Hot'), and their interaction as fixed factors. Lizard identity was employed as a random intercept, and as a random slope we included the trial within each level of lizard identity. We used the posterior distributions of these models to evaluate differences in learning rate between treatments and species. Learning slopes were obtained using the estimates of the variable trial per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not. We used the pmcmc method to test whether those slopes or the comparissons between treatments (e.g. slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) or species were different from zero (two-tailed tests). We considered statistical significance if p-value < 0.05.

@@

**Our results suggest that prenatal CORT and temperature do not affect the behavioural flexibility of _L. delicata_ and _L. guichenoti_. We did not find any significant differences between treatments in the reversal task, and when we compared the estimated slopes between species. This is in contrast with our predictions, as we expected that _L. guichenoti_ would be more impacted by the prenatal environment than _L. delicata_. We also expected that prenatal CORT would impair the performance of the lizards in the task, and that higher temperatures would improve it. However, we did not find any evidence of this.

They learn on average, as shown by positive slopes that do not vary when only those "superperformers" are employed in the analyses.

Why no effect of treatment? 
    What is expected from CORT? Other studeis showing no effects or the opposite: @szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016. Multiple factors may affect this results, including the dose of CORT, the time of exposure, and the species. Reversal?? Check @lui2017chronic and @jonson1976effect.

    What is expected from temperature? Other studies showing no effects or the opposite: @abayarathna_effects_2020; @dayananda. But these studies are in other species and with temperatures above the natural range. None of them reversal?? Only reversal done with criterion, so cold were not tested

    Selective sieve? No effects because is too simple?

Why no differences between species?
    Other papers finding no effects in learning: @bezzina2014does. Here they did not learn at all, too complex? Related with absence of treatment effects?

    No differences in treatment effects

Conclusions (sum up the discussion)**

>>>>

Behavioural flexibility, i.e. the ability to adjust the behaviour adaptively in response to internal or external changes by problem-solving, innovation, or alterations in one behaviour in favour of another [@brown2010behavioral], is considered crucial for invasion success [@wright2010behavioral; @chapple2012can; @szabo2020can]. In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats, as behaving flexibly enhances their capacity to exploit new resources and face new challenges [@wright2010behavioral; @chapple2012can; @szabo2020can]. For instance, when closely related species of birds were compared, successful invaders showed a higher frequency of foraging innovations [@sol2000behavioural]. According to the adaptive flexibility hypothesis, behavioural flexibility can be more relevant during introduction, establishment, and spread of the population [@wright2010behavioral; @chapple2012can; @szabo2020can], stages that, nonetheless, can involve more than one generation. If behavioural flexibility is adaptive in these different stages, both this trait and the underlining mechanisms are expected to persist through generations in succesful invaders. For example, in mammals, birds, reptiles and amphibians, relative brain size is a good predictor of behavioural flexibility and invasive success [@sol2000behavioural; @sol2002behavioural; @sol2008brain; @amiel_effects_2017]. Since the brain is especially sensitive to environmental changes during the early stages of development [@zhu_prenatal_2004] it is possible that the new prenatal environment faced by establishing populations can affect the behavioural flexibility of the new generations [see @jonson1976effect; @lui2017chronic], with a consequent impact on invasion success. Therefore, we consider important to investigate the effect of developmental conditions on behavioural flexibility and if it can predict the success of invasive species.  
In this scenario, we consider prenatal Glucocorticoids (GCs) and thermal environment to be salient factors in the study of behavioural flexibility within the context of invasion biology for several reasons. First, GCs are the main group of hormones related to animals’ response towards alterations in homeostasis [@sapolsky_how_2000] that may become relevant when facing new environments and their new thermal conditions [@Crino_2023]. Secondly, there is evidence that both developmental GCs and thermal environment can impact cognition in different taxa [@jonson1976effect; @zhu_prenatal_2004; @amiel_hotter_2012; @lui2017chronic]. For instance, young Sprague-Dawley rats treated with Corticosterone (CORT) - the principal GC in rodents, birds, and reptiles [@Crino_2023] - performed worse than controls in a reversal learning task [@lui2017chronic], while exposure of pregnant guinea pigs to higher temperatures resulted in significant impariments with offspring behavoural flexibility [@jonson1976effect]. In addition, the cognitive effects of both factors are triggered by similar physiological mechanisms intertwined with brain development [@jonson1976effect; @coomber_independent_1997; @zhu_prenatal_2004; @amiel_effects_2017], an aspect that is considered essential in succesful invaders [@sol2000behavioural; @sol2002behavioural; @sol2008brain; @amiel_effects_2017].  
Here, we studied the interactive contribution of prenatal temperature and CORT on behavioural flexibility in two species that exhibit distinct outcomes in their roles as invasive species, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L. delicata_ has been succesful in colonising different areas around the globe [@baker1979rainbow; @chapple2013divergent; @chapple2015deliinvLHI]. Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L. delicata_, and are likely key for its invassive success [@chapple_know_2011; @chapple2012can; @bezzina2014does]. However, differences between species in behavioural flexibility and how this trait is affected by the prenatal environment remains unknown. To investigate these effects, we first manipulated CORT levels and incubation temperature in eggs from both species, and then subjected the juveniles to a reversal learning task as a measure of behavioural flexibility [@brown2010behavioral; @szabo2020can]. We predicted that _L. delicata_ would, on average, perform more proficiently in the task that _L. guichenoti_. We also anticipated prenatal CORT and temperature to affect the behavioural flexibility of both species, and that _L. guichenoti_ would be more impacted by the prenatal environment than _L. delicata_. Regarding the directionality of the treatments, high prenatal temperatures are considered to improve performance in cognitive tasks in ectotherms [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014], so we expected those incubated at higher temperatures to perform better in the reversal task. On the other hand, we expected that prenatal CORT would impair the performance of the lizards in the task, as it has been shown in other species [@lemaire_prenatal_2000; @zhu_prenatal_2004; @lui2017chronic]. Finally, we expected an interaction between both treatments, with high temperatures mitigating the negative effects of prenatal CORT on the performance of the lizards in the task.

@@

We performed the analyses for both species separately. We fitted a total of two different Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022). We had per each model four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations. We used 'Choice' - whether the individual chose correct (1) or not (0) - as the response variable, and trial, hormone ('CORT' versus 'Control'), incubation temperature ('Cold' versus 'Hot'), and their interaction as fixed factors. Lizard identity was employed as a random intercept, and as a random slope we included the trial within each level of lizard identity. We used the posterior distributions of these models to evaluate differences in learning rate between treatments and species. Learning slopes were obtained using the estimates of the variable trial per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not. We used the pmcmc method to test whether those slopes or the comparissons between treatments (e.g. slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) or species were different from zero (two-tailed tests). We considered statistical significance if p-value < 0.05.

@@

The study embarked on unraveling the interactive effects of prenatal temperature and Glucocorticoids (GCs) on behavioral flexibility in two closely related species, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). Despite sharing life history traits, only _L. delicata_ has proven successful in colonizing various global regions, and we hypothesized this may be related with higher behavioural flexibility. Furthermore, we expected behavioural flexibility in this species to be less susceptible to the effect of early environment since this trait can be adaptive in several stages of invasion that can involve more than one generation. Nevertheless, we did not find any significant differences between treatments or species when we compared their performance in a reversal task. We found that, on average, all the lizards from the four treatments and the two species learned the reversal task - indicated by estimated slopes that are significantly different from cero - and that they learn, on average, at the same rate. Furthermore, these results were consistent when only those individuals that performed better in the previous associative task were included (see Supplementary material).  
Previous studies have shown that _L. delicata_ is more explorative than _L. guichenoti_, which can be advantageous in some facets of the colonization of new environments [@chapple_know_2011]. However, @bezzina2014does found no differences in learning between the two skinks in an associative learning task. Our results are consistent with this study, and suggest that behavioural flexibility does not differ in these species. In addition, we did not find any effect of the treatments on both species. Both results contrast our original predictions, and may indicate that other aspects of their biology are more relevant for _L. delicata_ success as invaders than behavioural flexibility. For instance, the mentioned differences in exploratory behaviour can be the key factor explaining why _L. delicata_ and not _L. guichenoti_ is better at colonising new areas, since exploring more can enhance the ability to effectively locate novel resources in _L. delicata_ [@chapple_know_2011].  
Our early environment manipulation did not affect the performance of the lizards in the task. We expected CORT to decrease reversal learning abilities and high temperatures to minimize these effects. However, our results show no significant differences between treatments in both species. These results contrast with previous studies that show impairments of high GCs levels in different types of learning tests [@lemaire_prenatal_2000; @zhu_prenatal_2004; @farrell_developmental_2015-learn], but align with other experiments where GCs increased or did not impact learning abilities [@szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016]. More striking is the absence of effect of incubation temperature, since most of the studies in reptiles reported better performance when incubated at higher temperatures in different associative learning and motor tasks [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]; and those studies that reported cold-incubated lizards to learn faster [@dayananda_incubation_2017; @abayarathna_effects_2020] used incubation temepratures far beyond the natural range in the hot treatment. It could be arguable whether our results are caused by a low sample size or treatments (especially hormone) are not properly applied, but we find these possibilities unlikely for two main reasons. First, we have seen the effect of the treatments in other traits of the lizards (data non published), and second, both effects are still non significant when the data is pooled. In contrast, we find more plausible that the effects of the early environment can be task-dependent, as some studies show that the effects of prenatal GCs and temeprature vary depending on the region of the brain [@coomber_independent_1997; @sakata_neural_2000; @lemaire_prenatal_2000]. However, this is a speculation that needs to be tested in future studies that examine both the performance in different cognitive tasks as well as the neurophysiological effects.  
In conculsion, our results showed that lizards performed the reversal learning task - as measurement of behavioural flexibility - at a similar level, with no differences between treatments or species. These results contrast our original predictions, and may indicate that other aspects of their biology are more relevant for _L. delicata_ success as invaders than behavioural flexibility and the effects of developmental environment on this trait. However, we cannot discard that the effects of the early environment can be task-dependent, and future studies should examine both the performance in different cognitive tasks as well as the effect of both treatments in the brain.
